---
title: "Wine - Diagnostic Data Analytics"
author: "Jeffrey Graves"
date: "`r Sys.Date()`"
output: html_document
---


# Install and Load the Relevant Packages
```{r, message=FALSE}
#install.packages("ggplot2")
#install.packages("DT")
#install.packages("corrplot")
library("ggplot2")
library("DT")
library("corrplot")
```

# Import Data
```{r}
wine_df <- read.csv("./wine.csv", row.names=1)
red_df = wine_df[wine_df$color == "red",]
white_df = wine_df[wine_df$color == "white",]
```

# Investigating the Data
```{r, collapse=TRUE}
dim(wine_df)
datatable(wine_df, options=list(scrollX=TRUE))
```

# Diagnostic Analytics
## Variables
```{r}
variables <- colnames(wine_df)
numeric_vars <- variables[c(-1,-13)]
categorical_vars <- variables[c(1)]
```

## Kolmogorov-Smirnov Test
The Kolmogorov-Smirnov test can be used to test whether two samples come from the same distribution.
```{r, warning=FALSE}
p_values_df <- as.data.frame(sapply(
  numeric_vars,
  function(name) {
    ks.test(red_df[[name]], white_df[[name]])$p.value
  }
))
colnames(p_values_df) <- c("p-value")
datatable(p_values_df, options=list(scrollX=TRUE, pageLength=20))
```

## Correlation Analysis
### Pearson’s Correlation Coefficient
#### Overall
```{r}
correlations <- cor(wine_df[numeric_vars])
corrplot(
  correlations,
  type="upper",
  tl.col="black",
  tl.srt=45
)
corrplot(
  correlations,
  method="number",
  tl.col="black",
  tl.srt=45,
  number.cex=0.75
)
```

#### Red
```{r}
correlations <- cor(red_df[numeric_vars])
corrplot(
  correlations,
  type="upper",
  tl.col="black",
  tl.srt=45
)
corrplot(
  correlations,
  method="number",
  tl.col="black",
  tl.srt=45,
  number.cex=0.75
)
```

#### White
```{r}
correlations <- cor(white_df[numeric_vars])
corrplot(
  correlations,
  type="upper",
  tl.col="black",
  tl.srt=45
)
corrplot(
  correlations,
  method="number",
  tl.col="black",
  tl.srt=45,
  number.cex=0.75
)
```

### Spearman’s Rank Correlation Coefficient
#### Overall
```{r}
correlations <- cor(
  wine_df["quality"],
  wine_df[numeric_vars],
  method="spearman"
)
corrplot(
  correlations,
  type="upper",
  tl.col="black",
  tl.srt=45
)
corrplot(
  correlations,
  method="number",
  tl.col="black",
  tl.srt=45,
  number.cex=0.75
)
```

#### Red
```{r}
correlations <- cor(
  red_df["quality"],
  red_df[numeric_vars],
  method="spearman"
)
corrplot(
  correlations,
  type="upper",
  tl.col="black",
  tl.srt=45
)
corrplot(
  correlations,
  method="number",
  tl.col="black",
  tl.srt=45,
  number.cex=0.75
)
```

#### White
```{r}
correlations <- cor(
  white_df["quality"],
  white_df[numeric_vars],
  method="spearman"
)
corrplot(
  correlations,
  type="upper",
  tl.col="black",
  tl.srt=45
)
corrplot(
  correlations,
  method="number",
  tl.col="black",
  tl.srt=45,
  number.cex=0.75
)
```

### Scatter Plots
#### Overall
```{r}
pairs(
  wine_df[numeric_vars],
  pch=16,
  cex=0.1
)
```

#### Red
```{r}
pairs(
  red_df[numeric_vars],
  pch=16,
  cex=0.1
)
```

#### White
```{r}
pairs(
  white_df[numeric_vars],
  pch=16,
  cex=0.1
)
```