---
title: "Human Stress Detection in and through Sleep"
subtitle: "Team Maximus"
author: "Harrison Peloquin, Hunter Sawyer, Chris Dinkins, Jackson Hanchek"
date: "`r Sys.Date()`"
output:
  pdf_document: default
  html_document: default
---

# Library Installations  

```{r,echo=FALSE,message=FALSE,error=FALSE,warning=FALSE}
library(pander)
sleep_df <- read.csv("./SaYoPillow.xls", row.names=NULL)
```


# Explanation of Data source  

Our data comes from a sleep analysis conducted using a "smart-yoga pillow".  
the study studied 630 participants who slept on the pillow and noted a number of factors associated with each participant  
these factors are as follows 

* Snoring Range
* respiration  rate
* Body Temperature
* Limb Movement Rates
* Blood Oxygen Levels
* Eye movement Levels
* Number of Hours Slept
* Heart rate
* stress levels
  * based on a ranking 0-4 the next day
```{r,}

```

# Initial Observations

One initial thing to note about our data is that a minority of participants were ranked as having a hours slept value of 0.  
We decided that we would separate these participants into a separate data set because we were unsure how to interpret  
this value and we were worried it would influence the rest of the data.


# What we want to do with the data

  





## Citations


  L. Rachakonda, A. K. Bapatla, S. P. Mohanty, and E. Kougianos, “SaYoPillow: Blockchain-Integrated Privacy-Assured IoMT Framework for 
  Stress Management Considering Sleeping Habits”, IEEE Transactions on Consumer Electronics (TCE), Vol. 67, No. 1, Feb 2021, pp. 20-29. 
    
  L. Rachakonda, S. P. Mohanty, E. Kougianos, K. Karunakaran, and M. Ganapathiraju, “Smart-Pillow: An IoT based Device for Stress Detection     Considering Sleeping Habits”, in Proceedings of the 4th IEEE International Symposium on Smart Electronic Systems (iSES), 2018, pp. 161--166.  
